---
title: "Lasso regression for identify important attributes governing glass formation"
author: "Rongrong Dai"
date: "April 25, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the package and data

```{r message = FALSE, warning=FALSE}
library(dplyr)
library(glmnet)

data = read.csv("attributes_gfa_set AM to 1.csv", header = TRUE, sep = ",")
data$Class = as.factor(data$Class)
x = model.matrix(Class~.,data)[,-1]

y = data$Class
train = sample(1:nrow(x), nrow(x)*0.9) ##default replace = FALSE

test = (-train)

y.test = y[test]

```

## Fit a lasso model

```{r}
grid = 10^seq(10,-2,length=100)  ##a descending sequence 

lasso.mod = glmnet(x[train,],y[train],family="binomial", alpha = 1,lambda = grid)

```

## Use cross validation to choose the best lambda
```{r}

cv.out = cv.glmnet(x[train,],y[train],alpha=1, family="binomial")

par(mar=c(1,1,1,1))

bestlam = cv.out$lambda.min

print(paste("Best lambda = ", bestlam), sep="")

```

## Calculate the prediction accuracy, precision, and recall for the test data using Lasso

```{r}
lasso.pred=predict(lasso.mod,s=bestlam,newx=x[test,], family="binomial", type="class")
acc.table <- table(lasso.pred,y.test)
Accuracy <- round((acc.table[1,1] + acc.table[2,2])/sum(acc.table),4)
Precision <- round(acc.table[2,2]/(acc.table[2,2]+acc.table[2,1]),4)
Recall <- round(acc.table[2,2]/(acc.table[2,2]+acc.table[1,2]),4)
print(paste("Accuracy = ", Accuracy), sep="")
print(paste("Precision = ", Precision), sep="")
print(paste("Recall = ", Recall), sep="")

```

## Predictor selection: get the important attributes from absolute values of coefficients of the attributes

We can get the important attributes from absolute values of coefficients of the attributes. The table below print the attributes and their coffeicients arranged in descending order according to their importance. 

```{r}
out=glmnet(x,y,alpha=1,lambda=grid, family="binomial")

lasso.coef = predict(out,type="coefficients",s=bestlam, family="binomial")

index <- sapply(lasso.coef@i, function(x) x+1)

coeffi_table <- data.frame("var" = lasso.coef@Dimnames[[1]][index],"coef" = lasso.coef@x) %>% filter(var != "(Intercept)") %>% arrange(desc(abs(coef)))

coeffi_table
```

